Wine Quality Reds Exploration by HanByul Yang
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(knitr)
library(dplyr)
library(gridExtra)
library(psych)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
df <- read.csv('wineQualityWhites.csv')

# Remove unnecessary column
drops <- c("X")
wineQuality <- df[, !(names(df) %in% drops)]
```

# Univariate Plots Section
```{r echo=FALSE, Summary}
# Summarize the data set
dim(wineQuality)
names(wineQuality)
str(wineQuality)
summary(wineQuality)
```

The median qulity of red wine is 6. median residual.sugar is 2.2 g / dm^3.
median alcohol of red wine is 10.2 %.


```{r echo=FALSE, fixed.acidity}
# qplot(fixed.acidity, data = wineQuality)
qplot(fixed.acidity, data = wineQuality)
qplot(fixed.acidity, data = wineQuality, binwidth=0.5)
```

Most red wines have fixed acidity between 7.10 g/dm^3 and  9.20 g/dm^3.

```{r echo=FALSE, volatile.acidity}
qplot(volatile.acidity, data = wineQuality, binwidth = 0.01)
qplot(volatile.acidity, data = wineQuality, binwidth = 0.01) +
  coord_cartesian(xlim=c(0, 1.1))
```

Most red wines have volatile acidity between 4 g/dm^3 and 7 g/dm^3. There are some outliers.

```{r echo=FALSE, citric.acid}
qplot(citric.acid, data = wineQuality, binwidth = 0.01)
qplot(citric.acid, data = wineQuality, binwidth = 0.01) + 
  coord_cartesian(xlim=c(0, 1.1))
table(wineQuality$citric.acid == 0)
```

138 red wines have 0 g/dm^3 citric acid. There is an outlier that has 1.0 g/dm^3.

```{r echo=FALSE, table_citric.acid}
summary(wineQuality$citric.acid)
sort(table(wineQuality$citric.acid), decreasing = T)
```

```{r echo=FALSE, residual.sugar}
qplot(residual.sugar, data = wineQuality)
qplot(residual.sugar, data = wineQuality, binwidth = 0.1) + 
  coord_cartesian(xlim=c(0, 40))
qplot(residual.sugar, data = wineQuality, binwidth = 0.05) + 
  scale_x_log10()
summary(wineQuality$residual.sugar)
```

Most of red wines have residual sugar between 1.9 g/dm^3 to 2.6 g/dm^3: median 2.2g/dm^3 and mean 2.539 g/dm^3.

```{r echo=FALSE, chlorides}
qplot(chlorides, data = wineQuality, binwidth = 0.005)
qplot(chlorides, data = wineQuality, binwidth = 0.005) + 
  coord_cartesian(xlim=c(0, 0.25))
qplot(chlorides, data = wineQuality, binwidth = 0.01) +
  scale_x_log10(breaks = c(0.01, 0.025, 0.05, 0.1, 0.2, 0.5))
summary(wineQuality$chlorides)
```

Most of red wines have chlorides between 0.07 g/dm^3 to 0.09 g/dm^3: median 0.079 g/dm^3 and mean 0.08747 g/dm^3.

```{r echo=FALSE, table_chlorides}
sort(table(wineQuality$chlorides), decreasing = T)
```

```{r echo=FALSE, free.sulfur.dioxide}
qplot(free.sulfur.dioxide, data = wineQuality)
qplot(free.sulfur.dioxide, data = wineQuality, binwidth = 0.5)
qplot(free.sulfur.dioxide, data = wineQuality, binwidth = 0.1) +
  scale_x_log10()
summary(wineQuality$free.sulfur.dioxide)
```

Most free.sulfur.dioxide values are integers. Most red wines have a free.sulfur.dioxide between 7 mg/dm^3 and 21 mg/dm^3.

```{r echo=FALSE, total.sulfur.dioxide}
qplot(total.sulfur.dioxide, data = wineQuality, binwidth = 5)
qplot(total.sulfur.dioxide, data = wineQuality, binwidth = 5) + 
  coord_cartesian(xlim=c(0, 300))
summary(wineQuality$total.sulfur.dioxide)
```
All of total.sulfur.dioxide values are integers. Most red wines have a total.sulfur.dioxide between 22 mg/dm^3 and 62 mg/dm^3. There are some outliers.

```{r echo=FALSE, density}
qplot(density, data = wineQuality, binwidth = 0.0005)
qplot(density, data = wineQuality, binwidth = 0.0005) +
  coord_cartesian(xlim=c(0.985, 1.005))
summary(wineQuality$density)
```

```{r echo=FALSE, pH}
qplot(pH, data = wineQuality, binwidth = 0.05)
summary(wineQuality$pH)
```

Most of red wines have a pH between 3.21 and 3.4: median 3.31 and mean 3.311.

```{r echo=FALSE, sulphates}
qplot(sulphates, data = wineQuality, binwidth = 0.05)
qplot(sulphates, data = wineQuality, binwidth = 0.05) +
  coord_cartesian(xlim=c(0.15, 1.2))
summary(wineQuality$sulphates)
```

```{r echo=FALSE, alcohol}
qplot(alcohol, data = wineQuality, binwidth = 0.1)
summary(wineQuality$alcohol)
```

Most of red wines have a alcohol between 9.5 and 11.1: median 10.2 and mean 10.42.

```{r echo=FALSE, quality}
qplot(x = quality, data = wineQuality)
summary(wineQuality$quality)
```

All of quality values are integers and between 3 and 8.
Most of red wines have a quality between 5 and 6: median 6 and mean 5.636

```{r echo=FALSE, table_quality}
sort(table(wineQuality$quality), decreasing = T)
```

```{r echo=FALSE, factor_quality}

# attach(wineQuality)
# wineQuality$quality_class[quality <= 3] <- 1
# wineQuality$quality_class[quality > 3 & quality <= 6] <- 2
# wineQuality$quality_class[quality > 6] <- 3
# detach(wineQuality)
# summary(wineQuality$quality_class)
# 
# wineQuality$quality_class <- factor(wineQuality$quality_class, labels = c("low", "mid", "high"))
wineQuality$quality_class <- factor(wineQuality$quality, labels = c(3:9))
levels(wineQuality$quality_class)
summary(wineQuality$quality_class)
```

# Univariate Analysis

### What is the structure of your dataset?

### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?



# Bivariate Plots Section
```{r echo=FALSE, Correlation_matrix}
cor(wineQuality[, c(1:12)])
```

The alcohol and density are the most correlated with quality.

```{r, echo=FALSE, Pairs_Panel}
pairs.panels(wineQuality[, c(1:12)])
```

```{r echo=FALSE, quality_vs_alcohol}
qplot(y = alcohol, x = quality, data = wineQuality)
```

```{r, echo=FALSE, quality_vs_density}
qplot(y = density, x = quality, data = wineQuality)
qplot(y = density, x = quality, data = wineQuality) +
  coord_cartesian(ylim = c(0.985, 1.005))
```

```{r, echo=FALSE, quality_class}
qplot(quality_class, data = wineQuality, geom="bar")
```

```{r, echo=FALSE, quality_class_vs_alcohol}
qplot(x = quality_class, y = alcohol, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = alcohol, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim=c(8, 13))
```

```{r, echo=FALSE, density_vs_alcohol}
qplot(x = quality_class, y = density, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = density, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(0.985,1.005))
```

```{r, echo=FALSE, quality_class_vs_fixed.acidity}
qplot(x = quality_class, y = fixed.acidity, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = fixed.acidity, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(5,11))
```

```{r, echo=FALSE, quality_class_vs_volatile.acidity}
qplot(x = quality_class, y = volatile.acidity, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = volatile.acidity, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(0.15,0.5))
```

```{r, echo=FALSE, quality_class_vs_citric.acid}
qplot(x = quality_class, y = citric.acid, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = citric.acid, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(0.0, 0.7))
```

```{r, echo=FALSE, quality_class_vs_residual.sugar}
qplot(x = quality_class, y = residual.sugar, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = residual.sugar, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(0, 25))
```

```{r, echo=FALSE, quality_class_vs_chlorides}
qplot(x = quality_class, y = chlorides, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = chlorides, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(0, 0.09))
```

```{r, echo=FALSE, quality_class_vs_free.sulfur.dioxide}
qplot(x = quality_class, y = free.sulfur.dioxide, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = free.sulfur.dioxide, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(0, 90))
```

```{r, echo=FALSE, quality_class_vs_total.sulfur.dioxide}
qplot(x = quality_class, y = total.sulfur.dioxide, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = total.sulfur.dioxide, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(0, 300))
```

```{r, echo=FALSE, quality_class_vs_pH}
qplot(x = quality_class, y = pH, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = pH, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(2.7, 3.7))
```

```{r, echo=FALSE, quality_class_vs_sulphates}
qplot(x = quality_class, y = sulphates, data = wineQuality, geom="boxplot")
qplot(x = quality_class, y = sulphates, data = wineQuality, geom="boxplot") +
  coord_cartesian(ylim = c(0.2, 0.9))
```
```{r, echo=FALSE, quality_vs_alcohol_linear_model}
quantile(wineQuality$alcohol, probs = seq(0, 1, 0.01))
quantile(wineQuality$alcohol, 0.999)
qplot(x = alcohol, y = quality, color = I('orange'), alpha = I(1/20),
      data = subset(wineQuality,
                    alcohol > 0 & alcohol <= quantile(wineQuality$alcohol, 0.999))) +
  geom_smooth(method = 'lm', se = T, color = 'blue')

```

```{r echo=FALSE, Linear_Model}
alcoholQualityLM <- lm(quality ~ alcohol,
                    data = subset(wineQuality,
                                  alcohol > 0 & alcohol <= quantile(wineQuality$alcohol, 0.999)))
summary(alcoholQualityLM)
```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}
makeProfilePlot <- function(mylist,names)
{
     require(RColorBrewer)
     # find out how many variables we want to include
     numvariables <- length(mylist)
     # choose 'numvariables' random colours
     colours <- brewer.pal(numvariables,"Set1")
     # find out the minimum and maximum values of the variables:
     mymin <- 1e+20
     mymax <- 1e-20
     for (i in 1:numvariables)
     {
        vectori <- mylist[[i]]
        mini <- min(vectori)
        maxi <- max(vectori)
        if (mini < mymin) { mymin <- mini }
        if (maxi > mymax) { mymax <- maxi }
     }
     # plot the variables
     for (i in 1:numvariables)
     {
        vectori <- mylist[[i]]
        namei <- names[i]
        colouri <- colours[i]
        if (i == 1) { plot(vectori,col=colouri,type="l",ylim=c(mymin,mymax)) }
        else         { points(vectori, col=colouri,type="l")                                     }
        lastxval <- length(vectori)
        lastyval <- vectori[length(vectori)]
        text((lastxval-10),(lastyval),namei,col="black",cex=0.6)
     }
 }

library(RColorBrewer)
names <- c("alcohol","density","residual.sugar","sulphates","pH")
mylist <- list(wineQuality$alcohol,wineQuality$density,wineQuality$residual.sugar,wineQuality$sulphates,wineQuality$pH)
makeProfilePlot(mylist,names)
```

```{r}
sapply(wineQuality[1:12],mean)
```

```{r}
printMeanAndSdByGroup <- function(variables,groupvariable)
{
     # find the names of the variables
     variablenames <- c(names(groupvariable),names(as.data.frame(variables)))
     # within each group, find the mean of each variable
     groupvariable <- groupvariable[,1] # ensures groupvariable is not a list
     means <- aggregate(as.matrix(variables) ~ groupvariable, FUN = mean)
     names(means) <- variablenames
     print(paste("Means:"))
     print(means)
     # within each group, find the standard deviation of each variable:
     sds <- aggregate(as.matrix(variables) ~ groupvariable, FUN = sd)
     names(sds) <- variablenames
     print(paste("Standard deviations:"))
     print(sds)
     # within each group, find the number of samples:
     samplesizes <- aggregate(as.matrix(variables) ~ groupvariable, FUN = length)
     names(samplesizes) <- variablenames
     print(paste("Sample sizes:"))
     print(samplesizes)
}
printMeanAndSdByGroup(wineQuality[1:12],wineQuality[12])
```

```{r}
calcWithinGroupsVariance <- function(variable,groupvariable)
  {
     # find out how many values the group variable can take
     groupvariable2 <- as.factor(groupvariable[[1]])
     levels <- levels(groupvariable2)
     numlevels <- length(levels)
     # get the mean and standard deviation for each group:
     numtotal <- 0
     denomtotal <- 0
     for (i in 1:numlevels)
     {
        leveli <- levels[i]
        levelidata <- variable[groupvariable==leveli,]
        levelilength <- length(levelidata)
        # get the standard deviation for group i:
        sdi <- sd(levelidata)
        numi <- (levelilength - 1)*(sdi * sdi)
        denomi <- levelilength
        numtotal <- numtotal + numi
        denomtotal <- denomtotal + denomi
     }
     # calculate the within-groups variance
     Vw <- numtotal / (denomtotal - numlevels)
     return(Vw)
  }
```

```{r}
cor.test(wineQuality$quality, wineQuality$alcohol)
```

```{r}
mosthighlycorrelated <- function(mydataframe,numtoreport)
  {
     # find the correlations
     cormatrix <- cor(mydataframe)
     # set the correlations on the diagonal or lower triangle to zero,
     # so they will not be reported as the highest ones:
     diag(cormatrix) <- 0
     cormatrix[lower.tri(cormatrix)] <- 0
     # flatten the matrix into a dataframe for easy sorting
     fm <- as.data.frame(as.table(cormatrix))
     # assign human-friendly names
     names(fm) <- c("First.Variable", "Second.Variable","Correlation")
     # sort and print the top n correlations
     head(fm[order(abs(fm$Correlation),decreasing=T),],n=numtoreport)
}
mosthighlycorrelated(wineQuality[1:12], 10)
```

```{r}
standardisedconcentrations <- as.data.frame(scale(wineQuality[1:12]))
# sapply(standardisedconcentrations,mean)
# sapply(standardisedconcentrations,sd)
wineQuality.pca <- prcomp(standardisedconcentrations)
summary(wineQuality.pca)
screeplot(wineQuality.pca, type="lines")
wineQuality.pca$rotation[,1]
```


```{r echo=FALSE, Build_Linear_Model}
m1 <- lm(quality ~ alcohol,
                    data = subset(wineQuality,
                                  alcohol > 0 & alcohol <= quantile(wineQuality$alcohol, 0.999)))
m2 <- update(m1, ~ . + density)
m3 <- update(m2, ~ . + sulphates)
mtable(m1, m2, m3)
```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
